### 2024-01-21 outputs/paper/paper.qmd
paper code 1/20

---
title: "Data Analysis of Mental Health Situation from 2014 to 2023 in Toronto"
title-block-banner: true
author: "Mary Cheng"
thanks: "Code and data are avaliable at:https://github.com/marycx/Analysis-of-Mental-Health-Situation-from-2014-to-2023-in-Toronto.git"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "something here"
format: pdf
toc: true
number-sections: true
---

```{r}
#| message: false
#| echo: false

#### Workspace set-up ####
install.packages("tidyverse")
install.packages("knitr")
install.packages("kableExtra")
install.packages("here")
install.packages("ggbeeswarm")
library(tidyverse) # a collection of data-related packages
library(knitr) # for kable
library(kableExtra)
library(here)
library(ggbeeswarm)
```

##