### 2024-01-21 outputs/paper/paper.qmd
update paper 1/21

---
title: "Data Analysis of Mental Health Situation from 2010 to 2017 in Toronto"
title-block-banner: true
author: "Mary Cheng"
thanks: "Code and data are avaliable at:https://github.com/marycx/Analysis-of-Mental-Health-Situation-from-2014-to-2023-in-Toronto.git"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "something here"
format: pdf
toc: true
number-sections: true
---

```{r}
#| message: false
#| echo: false
#### Workspace set-up ####
library(tidyverse) # a collection of data-related packages
library(knitr) # for kable
library(kableExtra)
library(here)
library(ggbeeswarm)
```

```{r}
# load data
mental_health_data = read_csv(
  file = here("inputs/data/cleaned_data.csv"),
  show_col_types = FALSE
)
```

```{## Results}
```

```{r}
# Show a piece of sample data 
#| message: false
#| echo: false
#| label: tbl_sample_cleaned_data
#| tbl_cap: Sample of cleaned mental health data

mental_health_data |>
  slice(1:10) |>
  kable(
    col.names = c("Occurred Year", "Sex", "Age Group", "Premises Type"),
    digits = 1,
    booktabs = TRUE,
    linesep = "",
    align = c("c", "c", "c", "c")
  )

```

```{r}
#| echo: false
#| message: false
#| label: tbl-sum_occurences
#| tbl-cap: Number of mental health apprehensions in Toronto each year from 2015-2017

mental_health_data |>
  select(occurred_year) |>
  count(occurred_year) |>
  kable(
    col.names = c("Occurred Year", "Count"),
    booktabs = TRUE,
    align = c("c", "c")
  )
```

```{r}
# Analyze the relationship between mental health apprehension and gender
#| echo: false
#| message: false
#| label: fig-sex_year
#| fig-cap: Number of male and female mental health apprehensions in Toronto

mental_health_data |>
  ggplot(mapping = aes(x = occurred_year, fill = sex)) +
  geom_bar(position = "dodge2") +
  labs(x = "Occurred Year", y = "Number of Apprehensions", fill = "Sex") +
  theme(legend.position = "bottom")
```

```{r}
#| echo: false
#| message: false
#| label: tbl-sum_occurences
#| tbl-cap: Total number of mental health apprenhension in age groups in Toronto from 2015 - 0217

mental_health_data |>
  select(age_group) |>
  count(age_group) |>
  kable(
    col.names = c("Age Group", "Count"),
    booktabs = TRUE,
    align = c("c", "c")
  )
```

```{r}
# Analyze the relationship between mental health apprehension and age groups in total
#| echo: false
#| message: false
#| label: fig-age_year
#| fig-cap: Total number of mental health apprenhension in age groups

mental_health_data |> 
  count(age_group) |> 
  ggplot(mapping = aes(x = age_group, y = n)) +
  geom_col() +
  theme_minimal() +
  labs(x = "Age group", y = "Number of mental health apprenhensions")
```

```{r}
# Analyze the relationship between mental health apprehension and age groups in each year
#| echo: false
#| message: false
#| label: fig-age_year_each
#| fig-cap: Number of mental health apprenhension in age groups in each year

mental_health_data |>
  ggplot(mapping = aes(x = occurred_year, fill = age_group)) +
  geom_bar(position = "dodge2") +
  labs(x = "Year", y = "Number of mental health apprenhension", fill = "Age group") +
  theme(legend.position = "bottom")
```

```{r}
# Analyze the relationship between mental health apprehension and premises type in each year
#| echo: false
#| message: false
#| label: fig-premise_type
#| fig-cap: Percentage of mental helath apprehensions by premises type

mental_health_data |>
  count(premises_type) |>
  ggplot(mapping = aes(x="",y = n, fill = premises_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y") +
  theme_minimal() +
  labs(x = "", y = "Percentage of mental health apprenhension in each premises type", fill = "Age group") +
  geom_text(aes(label = sprintf("%.1f%%", n/nrow(mental_health_data)*100)),
            position = position_stack(vjust = 0.5), color = "white", size = 3)

```

##